Death By Dropdown?

### 2025-04-01 conference-materials/2025_shiny-conf/slides.qmd
Updates content for slides 7 and 11
- Standardizes color codes for consistency across slides.
- Refines text formatting and layout for improved readability.
- Enhances speaker notes by making psych concepts more intuitive and concise.

## The Problem {.center background-color="#4682B4"}

@@

::: {.r-fit-text}
By shifting the focus from <span style="color:#D55E00">**cramming data**</span> to <span style="color:#009E73">**crafting stories**</span>
:::

@@

## BID In Action: Dashboard Transformation {.center background-color="#F2F2F2"}

::: {.absolute top="80%" left="12%" width="80%"}
<!-- TODO: exchange for gif? -->
![](img/before-bid.png)
:::

@@

## Behavior Insight Design (BID) Framework {background-color="#E6F0FF"}

@@

## Where BID Fits In {background-color="#F2F2F2"}

@@

## SoundPulse Case Study: BID Stages 1-3 {background-color="#F9F9F9"}

:::: {.columns}
::: {.column width="50%"}
- **Stage 1: Notice**
  * Complex filtering overwhelmed users (Cognitive Load)
  * 72% users reported confusion (Hick's Law)

- **Stage 2: Interpret**
  * Intuitive visualizations (Processing Fluency)
  * Clearer key insights (Data Storytelling)
:::

::: {.column width="50%"}
- **Stage 3: Structure**
  * Group related elements (Principle of Proximity)
  * Prioritization of metrics (Dual-Processing Theory)

<!-- TODO: consider adding a visual? otherwise, clean up text more -->
<!-- ![](img/small-dashboard-snippet.png){width="80%"} -->
:::
::::

::: {.notes}
**Case Study Setup**: Establish context for upcoming code examples.

- "Let's see how we assessed some of the challenges SoundPulse faced and how we applied the first 3 stages of the BID framework to address them"
- For Stage 1 (Notice): "They identified how filtering complexity was causing cognitive overload"
- For Stage 2 (Interpret): "They discovered users needed to process key metrics instantly"
- For Stage 3 (Structure): "They needed to group related elements and establish clear visual hierarchy"
- "Now let's look at how we implemented these principles using {reactable}..."
:::

## {reactable} Implementation: Visual Result {background-color="#F9F9F9"}

@@

## Implementation Demo: {reactable} Basics {background-color="#F9F9F9"}

@@

  # Stage 2: Processing Fluency - Color coding for instant insights
  columns = list(
    Metric1 = colDef(
      style = function(value) {
        # Color coding enhances processing fluency
        if (value > 50) {
          list(background = "rgba(0, 128, 128, 0.2)", 
               color = "#006666", fontWeight = "bold")
        } else if (value < 30) {
          list(background = "rgba(255, 140, 0, 0.2)", 
               color = "#DD8500", fontWeight = "bold")
        }
      }
    )
  )
)
```

@@

## Implementation Demo: {reactable} Advanced {background-color="#F9F9F9"}

@@

## SoundPulse Case Study: BID Stages 4-5 {background-color="#F9F9F9"}

:::: {.columns}
::: {.column width="50%"}

- **Stage 4: Anticipate**
  * Executives comparing metrics without context (Anchoring Effect)
  * Teams interpreting data differently (Confirmation Bias)
  * Different perspectives needed for different teams (Framing & Loss Aversion)
:::

::: {.column width="50%"}
- **Stage 5: Validate**
  * Teams needed actionable summaries (Peak-End Rule)
  * No clear path to consensus (Cooperation & Coordination)

<!-- TODO: consider adding a visual? otherwise, reorganize text -->
<!-- ![](img/small-dashboard-snippet2.png){width="80%"} -->
:::
::::

::: {.notes}
**Advanced Stages Case Study**: Set up context for echarts4r examples.

- "Now we'll look at how we tackled the later stages of the BID framework in SoundPulse's dashboard..."
- For Stage 4 (Anticipate): "We needed to address how cognitive biases were affecting their interpretations of the data"
- For Stage 5 (Validate): "We needed to create lasting impressions and facilitate team consensus"
- "These stages address the subtler psychological aspects of data interpretation"
- "Now let's see how we implemented these principles using {echarts4r}"
:::

## {echarts4r} Implementation: Visual Result {background-color="#F9F9F9"}

@@

## Implementation Demo: {echarts4r} Basics {background-color="#F9F9F9"}

@@

## Implementation Demo: {echarts4r} Reference Lines {background-color="#F9F9F9"}

@@

## Implementation Demo: {echarts4r} Framing Toggle {background-color="#F9F9F9"}

@@

## Interactive Demonstration: Framing Effect {background-color="#E6F0FF"}

@@

## From Framework to Workflow: Introducing {bidux} {.center background-color="#E6F0FF"}

@@

## What {bidux} Looks Like Today {background-color="#F9F9F9"}

@@

## SoundPulse Results {.center background-color="#4682B4"}

@@

- These outcomes reflect consistent feedback across client organizations that have embraced our framework. The qualitative data tells a compelling story:
  * (PAUSE) Teams are making decisions with much greater speed and confidence
  * (PAUSE) Users are discovering more actionable insights in each session
  * (PAUSE) And perhaps most telling—stakeholders have shifted from being skeptical to becoming advocates
- We're currently building telemetry into our dashboards to gather more precise measurements...but the message from our users is already clear: this framework is changing how teams interact with their data
- When the testimonial appears: "This kind of transformation is what BID is all about"
- "Remember, this wasn't just visual changes - this was a fundamental shift in how our users...and developers...approached the dashboard"
:::

## Key Takeaways {.center background-color="#4682B4"}

@@

## Thank You! {.center background-color="#4682B4"}

>>>>

## The Problem {.center background-color="#4682b4"}

@@

::: {.r-fit-text}
By shifting the focus from <span style="color:#d55e00">**cramming data**</span> to <span style="color:#009e73">**crafting stories**</span>
:::

@@

## BID In Action: Dashboard Transformation {.center background-color="#f2f2f2"}

::: {.absolute top="80%" left="12%" width="80%"}
![](img/before-bid.png)
:::

@@

## Behavior Insight Design (BID) Framework {background-color="#e6f0ff"}

@@

## Where BID Fits In {background-color="#f2f2f2"}

@@

## SoundPulse Case Study: BID Stages 1-3 {background-color="#f9f9f9"}

:::: {.columns .nonincremental .r-fit-text}

::: {.column width="32%"}
- <span style="color:#1c6d7d">**Stage 1: Notice**</span>
  * Complex filtering overwhelmed users (*Cognitive Load*)
  * 72% users reported confusion (*Hick's Law*)
:::

::: {.column width="32%"}
- <span style="color:#35a4ae">**Stage 2: Interpret**</span>
  * Intuitive visualizations (*Processing Fluency*)
  * Clearer key insights<br>(*Data Storytelling*)
:::

::: {.column width="32%"}
- <span style="color:#ff9000">**Stage 3: Structure**</span>
  * Group related elements (*Principle of Proximity*)
  * Prioritization of metrics (*Dual-Processing Theory*)
:::

::: {.absolute bottom="15%" left="15%" width="70%" height="30%"}
![](img/bid-framework.png)
:::

::: {.notes}
**Case Study Introduction (1-1.5 min):**

- "So...let's take a look at how we used this framework to help SoundPulse refactor their dashboard."
- **Stage 1 (Notice):** The first thing we noticed...as I'm sure you all of you did as well...was that users were overwhelmed with options - dropdown menus and complex tables
  * This is what we call 'cognitive load' - when the brain gets overtaxed with information.
  * We also identified a violation of Hick's Law, which simply means that more choices lead to longer decision times and confusion.
- **Stage 2 (Interpret):** Once we identified these issues, we focused on something called 'Processing Fluency' - making information easier to digest.
  * This meant using color-coding and visual patterns that the brain can process instantly without conscious effort.
  * You'll see in our code examples in a moment just how we implemented this with color scales in reactable.
- **Stage 3 (Structure):** "We then restructured using the 'Principle of Proximity' - which is a fancy way of saying we started grouping related elements. 
  * Now there are a ton of ways you can do this
  * As you'll see in our code example, we redesigned the layout to visually group metrics that belong together, and highlight the metrics that mattered most to create a natural flow through the data
- Now you don't need to know all of these concepts to use the framework...we'll come back to that
  * Don't need to be a psychologist to use the framework
- But first Milena is going to show you *how* we translated these concepts into code...
:::

## {reactable} Implementation: Visual Result {background-color="#f9f9f9"}

@@

## Implementation Demo: {reactable} Basics {background-color="#f9f9f9"}

@@

  # Stage 2: Processing Fluency - Color coding for instant insights
  columns = list(
    Metric1 = colDef(
      style = function(value) {
        # Color coding enhances processing fluency
        if (value > 50) {
          list(background = "rgba(0, 128, 128, 0.2)", 
               color = "#006666", fontWeight = "bold")
        } else if (value < 30) {
          list(background = "rgba(255, 140, 0, 0.2)", 
               color = "#dd8500", fontWeight = "bold")
        }
      }
    )
  )
)
```

@@

## Implementation Demo: {reactable} Advanced {background-color="#f9f9f9"}

@@

## SoundPulse Case Study: BID Stages 4-5 {background-color="#f9f9f9"}

:::: {.columns .nonincremental .r-fit-text}

::: {.column width="48%"}
- <span style="color:#ffca90">**Stage 4: Anticipate**</span>
  * Executives comparing metrics without context (*Anchoring Effect*)
  * Teams interpreting data differently (*Confirmation Bias*)
  * Different perspectives needed for different teams (*Framing & Loss Aversion*)
:::

::: {.column width="48%"}
- <span style="color:#98d3ca">**Stage 5: Validate**</span>
  * Teams needed actionable summaries (*Peak-End Rule*)
  * No clear path to consensus<br>(*Cooperation & Coordination*)
:::
::::

::: {.absolute bottom="15%" left="15%" width="70%" height="30%"}
![](img/bid-framework.png)
:::

::: {.notes}
**Later stages:** Setup for echarts examples

- "For the later stages of the BID framework, we addressed more subtle psychological factors that affect how users interpret and act on data."
- **Stage 4 (Anticipate):** "We noticed executives were making flawed comparisons without proper context.
  * This relates to the 'Anchoring Effect' - where people rely too heavily on the first piece of information they see
  * As you'll see in our echarts4r example, we added dynamic benchmark lines to provide meaningful context
  * We also implemented a 'Framing Toggle' to address how different teams focus on different perspectives on the same data."
- **Stage 5 (Validate):** "Finally, we applied the 'Peak-End Rule' - the principle that people judge experiences mainly by their most intense point and how they end.
  * Our code example shows how we created summary cards that highlight key takeaways, ensuring users walk away with the most important insights."
- Let's take a closer look at that code now...
:::

## {echarts4r} Implementation: Visual Result {background-color="#f9f9f9"}

@@

## Implementation Demo: {echarts4r} Basics {background-color="#f9f9f9"}

@@

## Implementation Demo: {echarts4r} Reference Lines {background-color="#f9f9f9"}

@@

## Implementation Demo: {echarts4r} Framing Toggle {background-color="#f9f9f9"}

@@

## Interactive Demonstration: Framing Effect {background-color="#e6f0ff"}

@@

## From Framework to Workflow: Introducing {bidux} {.center background-color="#e6f0ff"}

@@

## What {bidux} Looks Like Today {background-color="#f9f9f9"}

@@

## SoundPulse Results {.center background-color="#4682b4"}

@@

- These outcomes reflect consistent feedback across client organizations that have embraced our framework. The qualitative data tells a compelling story:
  * (PAUSE) Teams are making decisions with much greater speed and confidence
  * (PAUSE) Users are discovering more actionable insights in each session
  * (PAUSE) And perhaps most telling—stakeholders have shifted from being skeptical to becoming advocates
- We're currently building telemetry into our dashboards to gather more precise measurements...but the message from our users is already clear: this framework is changing how teams interact with their data
- When the testimonial appears: "This kind of transformation is what BID is all about"
- "Remember, this wasn't just visual changes - this was a fundamental shift in how our users...and developers...approached the dashboard"

- "While we didn't collect formal metrics before and after implementation, the qualitative feedback was tremendous. SoundPulse reported dramatically faster decision-making, more insights per session, and a complete transformation in stakeholder satisfaction - from frustration to enthusiasm."
:::

## Key Takeaways {.center background-color="#4682b4"}

@@

## Thank You! {.center background-color="#4682b4"}
